\section{Algorithm Description}
\label{sec:algorithm}

In this section, we describe the basic backoff protocol from the perspective of a packet that arrives at some specific time. The algorithm has three parameters a joining probability $\start$, an update parameter $\update$ and a normalization factor $\norm$. On arrival, it initializes its \emph{joining probability} $p = \start$.

Initially, the packet is inactive.  It listens silently until one of two \emph{critical events} occurs: (1) it observes a silent round, or (2) it observes a decoding event.  In either case, it becomes active in the following round.

From this round in which it becomes active, the packet divides time into epochs.  An epoch ends (and the next epoch begins) when one of three conditions is satisfied: there is a silent round, there is a decoding event, or the epoch has reached $\threshold$ rounds.  Whenever an epoch ends, a new epoch begins in the next round.

When an epoch begins, a packet decides whether or not to \emph{join} the epoch: it chooses to join the epoch with probability $p$.  If it has joined the epoch, then it broadcasts in every round of the epoch until the epoch ends.

When an epoch ends, a packet updates its joining probability $p$, depending on how the epoch ended.  If the epoch ended because of a silent round, the transmission probability is updated to $p=p\cdot e^{\norm\update/\threshold}$. % OLD: $p = p(1 + 1/\threshold)$
 If the epoch ended because $\threshold$ rounds expired without a decoding event, then the probability is updated to $p=p\cdot e^{-\update}$. % OLD: $p = p/2$.  
 If the epoch ended with a successful decoding event, then $p$ remains unchanged.

When the packet is included in a successful decoding event, it has completed its mission and exits the system.

\begin{itemize}
    \item If packet $p$ is inactive:
    \begin{itemize}
        \item $p = p_0$.
        \item Wait until some round $t$ where a critical event occurs (i.e., round $t$ is silent or has a decoding event).
        \item Become active in round $t+1$ and start a new epoch.
    \end{itemize}
    \item When a new epoch begins:
    \begin{itemize}
        \item Join the epoch with probability $p$.
    \end{itemize}
    \item During an epoch, in round $t$:
    \begin{itemize}
        \item If $p$ has joined the epoch, then broadcast in round $t$.
        \item If round $t$ is silent, then end the epoch and set 
            $p = p\cdot e^{\norm\update/\threshold}\approx p\cdot\big(1+\frac{\norm\update}{\threshold}\big)$.
        \item If round $t$ contains a successful decoding event, the end the epoch.
        \item If round $t$ is the $\threshold$ round of the epoch, then end the epoch and set 
        $p = p\cdot e^{-\update}\approx p\cdot (1-\update)$.
        \item If the epoch has ended, begin a new epoch in round $t+1$.
    \end{itemize}
\end{itemize}

